Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-04-09 WSSSPE2_report/WSSSPE2_review.tex
Section 4: Touch up

Progress in scientific research is dependent on the quality and accessibility of
software at all levels. Hence it is critical to address challenges related to
the development, deployment, maintenance, and overall sustainability of reusable
software as well as education around software practices. These challenges can be
technological, policy based, organizational, and educational, and are of
interest to developers (the software community), users (science disciplines),
software-engineering researchers, and researchers studying the conduct of
science (science of team science, science of organizations, science of science
and innovation policy, and social science communities). The WSSSPE1 workshop
engaged the broad scientific community to identify challenges and best practices
in areas of interest to creating sustainable scientific software. WSSSPE2
invited the community to propose and discuss specific mechanisms to move towards
an imagined future practice for software development and usage in science and
engineering. The workshop included multiple mechanisms for participation,
encouraged team building around solutions, and identified risky solutions with
potentially transformative outcomes. It strongly encouraged participation of
early-career scientists, postdoctoral researchers, and graduate students.
\todo{record how/number the travel awards were created and recipients selected, etc?}

@@

\item \textbf{Robert R. Downs: Community Recommendations for Improving Sustainable
Scientific Software Practices.}
Robert R. Downs, of the Columbia University Center for International Earth
Science Information Network (CIESIN), presented Community Recommendations for
Improving Sustainable Scientific Software Practices (2014), and described a
focus group study that was conducted with other members of the Science Software
Cluster of the Federation of Earth Science Information Partners (ESIP),
including W. Christopher Lenhardt, Erin Robinson, Ethan Davis, and Nicholas
Weber. For the study, almost 300 attendees of the 2014 Summer ESIP Meeting were
invited to participate in simultaneous roundtable discussions, on the
sustainability of science software. Over two-thirds of the roundtable focus
groups provided responses to a semi-structured questionnaire that contained
three sets of questions. Initial analysis of the participants' responses to the
set of questions eliciting recommendations for near-term actions of the
community to improve sustainable software practices revealed several
suggestions. The recommendations for improving scientific software
sustainability included improving community engagement and collaborative
activities, increasing understanding and awareness, and creating incentives to
motivate sustainable science software practices. Based on the suggestions
offered by the focus group respondents for near-term actions, the ESIP Science
Software Cluster plans to engage the community to achieve progress in the near
future on the recommended activities for improving sustainable scientific
software practices.

Reference

Downs RR, Lenhardt WC, Robinson E, Davis E, Weber N. 2014. Community
Recommendations for Improving Sustainable Scientific Software Practices. 2nd
Workshop on Sustainable Software for Science: Practice and Experiences
(WSSSPE2), New Orleans, Louisiana, 16 November 2014.
http://wssspe.researchcomputing.org.uk/wp-content/uploads/2014/11/Downs.pdf and
http://dx.doi.org/10.7916/D8Q52NBC
\choinote{Choi to revise.}

>>>>

Progress in scientific research is dependent on the quality and accessibility of
software at all levels. Hence it is critical to address challenges related to
the development, deployment, maintenance, and overall sustainability of reusable
software as well as education around software practices. These challenges can be
technological, policy based, organizational, and educational, and are of
interest to developers (the software community), users (science disciplines),
software-engineering researchers, and researchers studying the conduct of
science (science of team science, science of organizations, science of science
and innovation policy, and social science communities). The WSSSPE1 workshop
engaged the broad scientific community to identify challenges and best practices
in areas of interest to creating sustainable scientific software. WSSSPE2
invited the community to propose and discuss specific mechanisms to move towards
an imagined future practice for software development and usage in science and
engineering. The workshop included multiple mechanisms for participation,
encouraged team building around solutions, and identified risky solutions with
potentially transformative outcomes. It strongly encouraged participation of
early-career scientists, postdoctoral researchers, and graduate students.
\todo{record how/number of  travel awards were created \& recipients selected?}

@@

\item \textbf{Robert R. Downs: Community Recommendations for Improving Sustainable
Scientific Software Practices.}
Robert R. Downs, of the Columbia University Center for International Earth
Science Information Network (CIESIN), 
%presented Community Recommendations for
%Improving Sustainable Scientific Software Practices (2014), and 
described a
focus group study that was conducted with W. Christopher Lenhardt, Erin Robinson, Ethan Davis, and Nicholas
Weber of the Science Software
Cluster of the Federation of Earth Science Information Partners (ESIP). For the study, almost 300 attendees of the 2014 Summer ESIP Meeting were
invited to participate in simultaneous roundtable discussions on  
sustainability of science software. Over two-thirds of the roundtable focus
groups   responded to a semi-structured survey that contained
three sets of questions eliciting recommendations for near-term actions of the
community to improve sustainable software practices. Initial analysis of the participants' responses to the questionnaire
 revealed several suggestions, which  included improving community engagement and collaborative
activities, increasing understanding and awareness, and creating incentives to
motivate sustainable science software practices. 
%Based on the recommendations
%offered by the focus group respondents for near-term actions, 
The ESIP Science
Software Cluster plans to engage the community to achieve progress in the near
future on the recommended activities for improving sustainable scientific
software practices.

 \begin{comment}
Downs RR, Lenhardt WC, Robinson E, Davis E, Weber N. 2014. Community
Recommendations for Improving Sustainable Scientific Software Practices. 2nd
Workshop on Sustainable Software for Science: Practice and Experiences
(WSSSPE2), New Orleans, Louisiana, 16 November 2014.
http://wssspe.researchcomputing.org.uk/wp-content/uploads/2014/11/Downs.pdf and
http://dx.doi.org/10.7916/D8Q52NBC
 \end{comment}
